Supporting Information: Time Series Modeling of Solute Transport in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2019-09-23 Ben_Manuscripts/stochastic_transport/Supporting_Information.tex
Figure generation scripts + figures

@@

  \subsection*{Autocorrelation of steps}

% From Sokolov paper: "Assigning different waiting times τ i to each step, and assuming that the
% steps are uncorrelated as in a regular RW, gives rise to the CTRW model" -- I
% might need to check autocorrelation of steps lengths when a hop occurs rather
% than every time step. Not sure if there is enough data for that, but could check 
% a particularly hoppy trajectory after longer simulation.

  Based on the previous two sections, our model can likey be studied as a CTRW. 
  However, it is still possible that our CTRW model might also be convoluted with
  an FBM or a RWF process. In a pure CTRW, the steps are uncorrelated. 
  Both FBM and RWF exhibit anti-correlated steps.

  The steps in our system are not correlated. We showed this by calculating the
  autocorrelation function (ACF) of the step lengths in the $z$-direction. The
  ACF of a representative trajectory is shown in Figure~\ref{fig:eth_autocorrelation}.

  \begin{figure}[!htb]
  \centering
% Generated with brownian_test.py -t PR_nojump.xtc -g PR.gro -r ETH (and appropriate uncommenting -- need to rework that script)
% in directory: /home/bcoscia/Documents/Gromacs/Transport/NaGA3C11/ETH/10wt
  \includegraphics[width=0.8\textwidth]{eth_autocorrelation.pdf}
  \caption{The autocorrelation function (right) of a representative ethanol
	   center of mass $z$-coordinate trajectory (left) almost immediately decays to zero,
	   indicating a complete loss of memory of it's previous position. Noise increases
	   at large time lags due to decreased sampling.}\label{fig:eth_autocorrelation}
  \end{figure}

@@

  In Figure~\ref{fig:hurst_autocorrelation}, we plotted Equation~\ref{eqn:fbm_autocorrelation}
  for different values of $H$. When $H > 0.5$, Equation~\ref{eqn:fbm_autocorrelation} decays
  slowly to zero meaning one needs to study large time lags with high frequency in order to
  obtain accurate estimates. Fortunately, all of our solutes show anti-correlated motion, so
  most of the information in Equation \ref{eqn:fbm_autocorrelation} is contained within the 
  first few lags.

@@

  The autocovariance function of Fractional L\'evy motion is different from fractional
  Brownian motion (see Equation~\ref{M-eqn:fbm_autocorrelation}
  and~\ref{M-eqn:flm_autocovariance} of the main text), but their autocorrelation 
  structures are the same. The autocovariance function of FLM is dependent on the 
  expected value of squared draws from the underlying L\'evy distribution, $E\big[L(1)^2\big]$. 
  This is effectively the distribution's variance, which is undefined for most 
  L\'evy stable distributions due to their heavy tails. As a consequence, one should 
  expect $E\big[L(1)^2\big]$ to grow as more samples are drawn from the distribution. 
  However, we are only interested in the autocorrelation function. In order to predict
  the Hurst parameter from the autocorrelation function, we must show that it has 
  well-defined structure and is independent of the coefficient in 
  Equation~\ref{M-eqn:flm_autocovariance} of the main text. In Figure~\ref{fig:flm_autocorrelation}, 
  we plot the average autocorrelation function from an FLM process with an increasing 
  number of observations per generated sequence. For all simulations we set $H$=0.35 
  and $\alpha$=1.4. The variance-normalized autocovariance function, i.e. the autocorrelation
  function does not change with increasing sequence length. Additionally, the
  autocorrelation function of FBM, with the same $H$ is the same.

  % supporting_figures/flm_autocov.py
  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{flm_autocovariance.pdf}
  \caption{The autocorrelation function of an FLM process does not change with 
  increasing sequence length (N). It shares the same autocorrelation function as
  fractional Brownian motion (FBM). All sequenced used to make this plot were 
  generated using $H$=0.35 and $\alpha$=1.4 (for FLM).}\label{fig:flm_autocovariance}
  \end{figure}

@@

  \subsection*{Achieving the right correlation structure}

@@

  Stoev and Taqqu note that the transition between negatively and positively correlated
  draws occurs when $H = 1/ \alpha$. When $\alpha=2$, the marginal distribution is 
  Gaussian and $H=0.5$ as expected from FBM. We corrected the input $H$ so that the 
  value of $H$ measured based on the output sequence equaled the desired $H$. We 
  first adjusted the value of $H$ by adding ($1 / \alpha - 0.5$), effectively 
  recentering the correlation sign transition for any value of $1 \leq \alpha \leq 2$.
  This correction alone does a good job for input $H$ values near 0.5, but is
  insufficient if one desires a low values of $H$. The exact correction to $H$ is 
  not obvious so we created a database of output $H$ values tabulated as a function
  of input $H$ and $\alpha$ values. Figure~\ref{fig:hurst_correction} demonstrates the
  results of applying our correction. Without the correction, FLM realizations are
  more negatively correlated. This would result in under-predicted mean squared
  displacements when applying the model.

@@

  \subsection*{FLM realizations from truncated L\'evy stable distributions}

  To generate realizations from an uncorrelated truncated L\'evy process, one would
  randomly sample from the base distribution and replace values that are too large
  with new random samples from the base distribution, repeating the process until
  all samples are under the desired cut-off.

  If one is to truncate an FLM process, one can apply the simple procedure above for 
  drawing uncorrelated values from the marginal L\'evy stable distribution, but after
  adding correlation, the maximum drawn value is typically lower than the limit set 
  by the user. Additionally, the shape of the distribution itself changes. Analogous
  to the database used to correct the Hurst parameter, we created a database to 
  correct the input truncation parameter (the maximum desired draw). The database
  returns the value of the truncation parameter that will properly truncate the
  output marginal distribution based on $H$, $\alpha$ and $\sigma$ (the width parameter).
  Figure~\ref{fig:truncation_correction} shows the result of applying our correction.

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{truncation_correction.pdf}
  \caption{We can accurately truncate the marginal distribution of FLM innovations by
  applying a correction to the input truncation parameter. We generated FLM sequences
  and truncated the initial L\'evy stable distribution (before Fourier transforming) at 
  a value of 0.5. After correlation structure is added, the width of the distribution 
  of fractional L\'evy noise decreases significantly. We corrected the input truncation
  parameter with our database resulting in a much more accurate distribution with 
  a maximum value close to 0.5.}\label{fig:truncation_correction}
  \end{figure}

>>>>

  \section{Solute Equilibration}\label{section:equilibration}

  We collected all data used for model generation after the solutes were 
  equilibrated. We assumed a solute to be equilibrated when the partition of
  solutes in and out of the pore region stopped changing. The pore region is
  defined as within 0.75 nm of the pore center. We've plotted the partition
  versus time in Figure~\ref{fig:equilibration} and indicated the chosen
  equilibration time points.

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{URE_equilibration.pdf}
  \caption{}\label{fig:URE_equilibration}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{GCL_equilibration.pdf}
  \caption{}\label{fig:GCL_equilibration}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{ACH_equilibration.pdf}
  \caption{}\label{fig:ACH_equilibration}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{MET_equilibration.pdf}
  \caption{}\label{fig:MET_equilibration}
  \end{subfigure}
  \caption{We considered a system to be equilibrated when the partition of solutes
  between the tails and pore plateaued. Our chosen equilibration point for each
  solute is indicated by the vertical black dashed line. (a) Urea equilibrates
  the fastest, after 1000 ns. (b) Ethylene glycol equilibrates after 1200 ns 
  (c) Acetic Acid does not equilibrate until 4400 ns. (d) We considered methanol
  to be equilibrated after 3500 ns. Methanol nearly completely partitions into
  the tails.}\label{fig:equilibration}
  \end{figure}

@@

@@

  In Figure~\ref{fig:hurst_autocorrelation}, we plotted Equation~\ref{eqn:fbm_autocorrelation}
  for different values of $H$. When $H > 0.5$, Equation~\ref{eqn:fbm_autocorrelation} decays
  slowly to zero meaning one needs to study large time lags with high frequency in order to
  accurately estimate $H$ from the data. Fortunately, all of our solutes show anti-correlated
  motion, so most of the information in Equation \ref{eqn:fbm_autocorrelation} is contained
  within the first few lags.

@@

  The autocovariance function of fractional L\'evy motion is different from fractional
  Brownian motion (see Equations~\ref{M-eqn:fbm_autocorrelation}
  and~\ref{M-eqn:flm_autocovariance} of the main text), but their autocorrelation 
  structures are the same. The autocovariance function of FLM is dependent on the 
  expected value of squared draws from the underlying L\'evy distribution, $E\big[L(1)^2\big]$. 
  This is effectively the distribution's variance, which is undefined for most 
  L\'evy stable distributions due to their heavy tails. As a consequence, one should 
  expect $E\big[L(1)^2\big]$ to grow as more samples are drawn from the distribution
  with the autocovariance function responding accordingly. 
  However, we are only interested in the autocorrelation function. In order to predict
  the Hurst parameter from the autocorrelation function, we must show that it has 
  a well-defined structure and is independent of the coefficient in 
  Equation~\ref{M-eqn:flm_autocovariance} of the main text. In Figure~\ref{fig:flm_autocorrelation}, 
  we plot the average autocorrelation function from an FLM process with an increasing 
  number of observations per generated sequence. For all simulations we set $H$=0.35 
  and $\alpha$=1.4. The variance-normalized autocovariance function, i.e. the autocorrelation
  function, does not change with increasing sequence length. Additionally, the
  autocorrelation function of FBM, with the same $H$, is the same.

  % supporting_figures/flm_autocov.py
  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{flm_autocovariance.pdf}
  \caption{The autocorrelation function of an FLM process does not change with 
  increasing sequence length (N). It shares the same autocorrelation function as
  fractional Brownian motion (FBM). All sequenced used to make this plot were 
  generated using $H$=0.35 and, for FLM, $\alpha$=1.4.}\label{fig:flm_autocorrelation}
  \end{figure}

@@

  \subsection{Truncated L\'evy stable hop distributions}\label{section:truncation}

  \textit{Determining where to truncate the hop distribution:} A pure
  L\'evy stable distribution has heavy tails which can lead to arbitrarily
  long hop lengths. Our distribution of hop lengths fits well to a L\'evy
  distribution near the mean, but under samples the tails. In 
  Figure~\ref{fig:truncation_correction} we compare the empirically 
  measured hop length distribution of Urea to its maximum likelihood fit to
  a L\'evy stable distribution. The ratio between the two distributions 
  at each bin is nearly 1 close to the center, indicating a near-perfect
  fit, larger than 1 slightly further from the center, suggesting that 
  we slightly over sample intermediate hop lengths, and below 1 far from 
  the center, indicating under sampling of extremely long hop lengths.
  Based on the plot, we chose a cut-off of 0.8 nm in order to compensate for
  over sampled intermediate hop lengths. We used the same procedure
  to determine truncation parameters for all solutes.

  \begin{figure}
  \centering
  \includegraphics[width=0.75\textwidth]{truncation_cutoff.pdf}
  \caption{The ratio between the empirical and maximum likelihood theoretical
  distribution quantifies the quality of fit as function of hop length. The fit
  is near-perfect close to the mean. Intermediate hop lengths are over sampled, 
  and the tails are under sampled. We used this type of plot to determine the
  appropriate place to truncate the L\'evy stable distributions.}\label{fig:truncation_cutoff}
  \end{figure}

  \textit{Generating FLM realizations from a truncated L\'evy distribution:}
  To generate realizations from an uncorrelated truncated L\'evy process, one would
  randomly sample from the base distribution and replace values that are too large
  with new random samples from the base distribution, repeating the process until
  all samples are under the desired cut-off.

  If one is to truncate an FLM process, one can apply the simple procedure above for 
  drawing uncorrelated values from the marginal L\'evy stable distribution, but after
  adding correlation, the maximum drawn value is typically lower than the limit set 
  by the user. Additionally, the shape of the distribution itself changes. Analogous
  to the database used to correct the Hurst parameter, we created a database to 
  correct the input truncation parameter (the maximum desired draw). The database
  returns the value of the truncation parameter that will properly truncate the
  output marginal distribution based on $H$, $\alpha$ and $\sigma$ (the width parameter).
  Figure~\ref{fig:truncation_correction} shows the result of applying our correction.
  Note that generating this database requires a significant amount of simulation and
  still likely doesn't perfectly correct the truncation parameter. The output leads
  to a somewhat fuzzy cut-off of the output distribution. This is likely beneficial
  since we observe a small proportion of hops longer the chosen truncation cut-off.

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{truncation_correction.pdf}
  \caption{We can accurately truncate the marginal distribution of FLM innovations by
  applying a correction to the input truncation parameter. We generated FLM sequences
  and truncated the initial L\'evy stable distribution (before Fourier transforming) at 
  a value of 0.5. After correlation structure is added, the width of the distribution 
  of fractional L\'evy noise decreases significantly. We corrected the input truncation
  parameter with our database resulting in a distribution close to the theoretical
  distribution (black dashed line) with a maximum value close to 0.5.}\label{fig:truncation_correction}
  \end{figure}

  \subsection{Achieving the right correlation structure}\label{section:flm_correlation}

@@

  Stoev and Taqqu note that the transition between negatively and positively correlated
  draws occurs when $H = 1/ \alpha$. When $\alpha=2$, the marginal distribution is 
  Gaussian and the transition occurs at $H=0.5$ as expected from FBM. We corrected 
  the input $H$ so that the value of $H$ measured based on the output sequence equaled
  the desired $H$. We first adjusted the value of $H$ by adding ($1 / \alpha - 0.5$),
  effectively recentering the correlation sign transition for any value of $1 \leq \alpha \leq 2$.
  This correction alone does a good job for input $H$ values near 0.5, but is
  insufficient if one desires a low value of $H$. The exact correction to $H$ is 
  not obvious so we created a database of output $H$ values tabulated as a function
  of input $H$ and $\alpha$ values. Figure~\ref{fig:hurst_correction} demonstrates the
  results of applying our correction. Without the correction, FLM realizations are
  more negatively correlated. This would result in under-predicted mean squared
  displacements when applying the model.

@@
